---
title: "Presenting Collated Ontology Terms - OMIM and FAERS Outcomes"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)

# Read environment file
#load_dot_env("config.env")
readRenviron(here("config.env"))

base_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

LLM_results_dir <- file.path(base_dir, "OMIM_FAERS_outcomes")

```

# Mapped Predicted and Observed Drug outcomes

**Predicted outcomes** - OMIM phenotpic outcomes resulting from genetic variation in drug target gene

**Observed outcomes** - FAERS adverse events associated with drugs. This dataset includes all outcomes and those that are pregnancy and congenitally related.

## The terms have been mapped in three steps:
    1. Biobert model to calculate cosine similarity between OMIM and FAERS terms. Filtered for the top 30 most similar FAERS terms for each OMIM term.
    2. Qwen2.5 model to generate a confidence score of 0-5 for each OMIM-FAERS term pair based on the closeness of the traits. Those with a score of 0 were filtered out.
    3. DeepSeek model to generate a confidence score of 0-5 for each OMIM-FAERS term pair based on the closeness of the traits. Those with a score 0-1 were rejected.

## Confidence Score Interpretation:
    5: Exact or near-perfect conceptual match
    4: Strong pathophysiological relationship (same disease category, direct complications)
    3: Moderate relationship (shared mechanisms, risk factors, or related conditions)
    2: Weak relationship (indirect connections, shared symptoms)
    1: Very loose association
    0: No association

# Drugs of Interest:
    * Captopril
    * Methotrexate
    * Mycophenolate mofetil
    * Thalidomide
    * Tretinoin
    * Valproate
    * Warfarin

# Captopril
## All outcomes

```{r}
#| label: Captopril all
#| include: true
#| echo: false

captopril_results_path <- file.path(LLM_results_dir, "captopril/all/captopril_omim_faers_LLM_all_accepted.csv")

captopril_results <- fread(captopril_results_path) %>%
    select(Predicted_term, Observed_term)

kable(captopril_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Captopril congenital
#| include: true
#| echo: false

captopril_results_path <- file.path(LLM_results_dir, "captopril/cong/captopril_omim_faers_LLM_cong_accepted.csv")

captopril_results <- fread(captopril_results_path) %>%
    select(Predicted_term, Observed_term)

kable(captopril_results) %>%
  kable_styling() 

```

# Methotrexate
## All outcomes

```{r}
#| label: Methotrexate all
#| include: true
#| echo: false

methotrexate_results_path <- file.path(LLM_results_dir, "methotrexate/all/methotrexate_omim_faers_LLM_all_accepted.csv")

methotrexate_results <- fread(methotrexate_results_path) %>%
    select(Predicted_term, Observed_term)

kable(methotrexate_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only
```{r}
#| label: Methotrexate cong
#| include: true
#| echo: false

methotrexate_results_path <- file.path(LLM_results_dir, "methotrexate/cong/methotrexate_omim_faers_LLM_cong_accepted.csv")

methotrexate_results <- fread(methotrexate_results_path) %>%
    select(Predicted_term, Observed_term)

kable(methotrexate_results) %>%
  kable_styling() 

```

# Mycophenolate mofetil
## All outcomes

```{r}
#| label: Mycophenolate mofetil all
#| include: true
#| echo: false

mycophenolate_mofetil_results_path <- file.path(LLM_results_dir, "mycophenolate_mofetil/all/mycophenolate_mofetil_omim_faers_LLM_all_accepted.csv")

mycophenolate_mofetil_results <- fread(mycophenolate_mofetil_results_path) %>%
    select(Predicted_term, Observed_term)

kable(mycophenolate_mofetil_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Mycophenolate mofetil cong
#| include: true
#| echo: false

mycophenolate_mofetil_results_path <- file.path(LLM_results_dir, "mycophenolate_mofetil/cong/mycophenolate_mofetil_omim_faers_LLM_cong_accepted.csv")

mycophenolate_mofetil_results <- fread(mycophenolate_mofetil_results_path) %>%
    select(Predicted_term, Observed_term)

kable(mycophenolate_mofetil_results) %>%
  kable_styling() 

```

# Thalidomide
## All outcomes

```{r}
#| label: Thalidomide all
#| include: true
#| echo: false

thalidomide_results_path <- file.path(LLM_results_dir, "thalidomide/all/thalidomide_omim_faers_LLM_all_accepted.csv")

thalidomide_results <- fread(thalidomide_results_path) %>%
     select(Predicted_term, Observed_term)

kable(thalidomide_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Thalidomide cong
#| include: true
#| echo: false

thalidomide_results_path <- file.path(LLM_results_dir, "thalidomide/cong/thalidomide_omim_faers_LLM_cong_accepted.csv")

thalidomide_results <- fread(thalidomide_results_path) %>%
     select(Predicted_term, Observed_term)

kable(thalidomide_results) %>%
  kable_styling() 

```

# Tretinoin
## All outcomes

```{r}
#| label: Tretinoin all
#| include: true
#| echo: false

tretinoin_results_path <- file.path(LLM_results_dir, "tretinoin/all/tretinoin_omim_faers_LLM_all_accepted.csv")

tretinoin_results <- fread(tretinoin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(tretinoin_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Tretinoin cong
#| include: true
#| echo: false

tretinoin_results_path <- file.path(LLM_results_dir, "tretinoin/cong/tretinoin_omim_faers_LLM_cong_accepted.csv")

tretinoin_results <- fread(tretinoin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(tretinoin_results) %>%
  kable_styling() 

```

# Valproate
## All outcomes

```{r}
#| label: Valproate all
#| include: true
#| echo: false

valproate_results_path <- file.path(LLM_results_dir, "valproate/all/valproate_omim_faers_LLM_all_accepted.csv")

valproate_results <- fread(valproate_results_path) %>%
     select(Predicted_term, Observed_term)

kable(valproate_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Valproate cong
#| include: true
#| echo: false

valproate_results_path <- file.path(LLM_results_dir, "valproate/cong/valproate_omim_faers_LLM_cong_accepted.csv")

valproate_results <- fread(valproate_results_path) %>%
     select(Predicted_term, Observed_term)

kable(valproate_results) %>%
  kable_styling() 

```

# Warfarin
## All outcomes

```{r}
#| label: Warfarin all
#| include: true
#| echo: false

warfarin_results_path <- file.path(LLM_results_dir, "warfarin/all/warfarin_omim_faers_LLM_all_accepted.csv")

warfarin_results <- fread(warfarin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(warfarin_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Warfarin cong
#| include: true
#| echo: false

warfarin_results_path <- file.path(LLM_results_dir, "warfarin/cong/warfarin_omim_faers_LLM_cong_accepted.csv")

warfarin_results <- fread(warfarin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(warfarin_results) %>%
  kable_styling() 

```

# Render
```{r}
quarto::quarto_render(
  "05_Ontology_Mapping/055b_Mapping_Results_Collation.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```