---
title: "MR Results - Drug Bank Primary Indications and openGWAS outcomes"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
        fig-width: 8
        fig-height: 10
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)
library(tidyr)
library(ieugwasr)
library(TwoSampleMR)
library(ggplot2)

# Read environment file

readRenviron(here("config.env"))

interim_dir <- Sys.getenv("interimdatadir")
processed_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

exposure_file <- file.path(interim_dir, "predicted_outcomes/Primary_Indication_Mapping/MR_input/exposure_associations.csv")
outcome_file <- file.path(interim_dir, "predicted_outcomes/Primary_Indication_Mapping/MR_input/outcome_associations.csv")
targets_file <- file.path(interim_dir, "predicted_outcomes/Drug_Bank_targets/Drug_Bank_targets.csv")
indications_file <- file.path(interim_dir, "reported_outcomes/Drug_Bank_outcomes/Drug_Bank_Indications.csv")
output_dir <- file.path(interim_dir, "predicted_outcomes/Primary_Indication_Mapping/MR_output")

mr_results_dir <- file.path(processed_dir,"Primary_indications_MR_Results")

```


```{r}
#| label: Reading in data
#| include: false

# Target of each drug
drug_targets <- fread(targets_file) %>% 
  filter(Primary == "Primary")

# Indication of each drug
drug_indications <- fread(indications_file) %>% 
  filter(`Approval Level` == "Prescription")

# Normalise some drug naming
drug_indications[which(drug_indications$Drug == "Valproate"), "Drug"] <- "Valproic acid"

# Drug target assocation data
exposure_data <- fread(exposure_file)

# Outcome association data
outcome_data <- fread(outcome_file) %>% 
  select("indication", "openGWAS_outcome", "id", "chr", "position", "rsid", 
         "ea", "nea", "eaf", "beta", "se", "p", "n")


```

# Drugs of Interest:
    * Captopril
    * Methotrexate
    * Mycophenolate mofetil
    * Thalidomide
    * Tretinoin
    * Valproate
    * Warfarin

# Captopril

## Primary Indications
```{r}
#| label: Captopril_Primary_Indications
#| include: true
#| echo: false

drug <- "Captopril"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Captopril Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Captopril openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Captopril_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Captopril Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Captopril_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Captopril_MR_Results
#| include: true
#| echo: false
#######################################################
# 11. Plotting
#######################################################

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```


# Methotrexate

## Primary Indications
```{r}
#| label: Methotrexate_Primary_Indications
#| include: true
#| echo: false

drug <- "Methotrexate"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Methotrexate Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Methotrexate openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Methotrexate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Methotrexate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Methotrexate_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Methotrexate_MR_Results
#| include: true
#| echo: false
#| fig-width: 10
#| fig-height: 15

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```

# Mycophenolate mofetil

## Primary Indications
According to DrugBank, Mycophenolate mofetil is used for the following indications:

 - Transplant organ rejection

However, it is reported as being used as Adjunct therapy, and has therefore beeen filtered out as not the primary drug used for this infication.
Therefore, there are no primary indications for this drug, and no associated outcomes in openGWAS. This is an example of a drug where the MR analysis is not informative, as the drug is not used as a primary treatment for any indication.


# Thalidomide

## Primary Indications

```{r}
#| label: Thalidomide_Primary_Indications
#| include: true
#| echo: false

drug <- "Thalidomide"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Thalidomide Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Thalidomide openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Thalidomide_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Thalidomide Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Thalidomide_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Thalidomide_MR_Results
#| include: true
#| echo: false
#| fig-width: 12
#| fig-height: 8

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```

# Tretinoin

## Primary Indications

```{r}
#| label: Tretinoin_Primary_Indications
#| include: true
#| echo: false

drug <- "Tretinoin"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Tretinoin Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Tretinoin openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Tretinoin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Tretinoin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Tretinoin_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Tretinoin_MR_Results
#| include: true
#| echo: false
#| fig-width: 10
#| fig-height: 10

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```

# Valproate

## Primary Indications

```{r}
#| label: Valproate_Primary_Indications
#| include: true
#| echo: false

drug <- "Valproic acid"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Valproate Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Valproate openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Valproate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Valproate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Valproate_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Valproate_MR_Results
#| include: true
#| echo: false
#| fig-width: 10
#| fig-height: 10

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```

# Warfarin

## Primary Indications

```{r}
#| label: Warfarin_Primary_Indications
#| include: true
#| echo: false

drug <- "Warfarin"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Warfarin Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Warfarin openGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Warfarin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Warfarin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis
```{r}
#| label: Warfarin_MR
#| include: false
#| echo: false

exposure_SNPs <- exposure_data[grep(target_list, exposure_data$drug_target)] %>% 
  as.data.frame()
  
# Extract all associated SNPs

SNP_list <- paste0("^", paste(unique(exposure_SNPs$rsid.x), collapse = "$|^"), "$")
  
# Extract association data for all SNPs and all available outcomes

outcome_SNPs <- outcome_data[grep(SNP_list, outcome_data$rsid),] %>% 
  as.data.frame()
  
# Format data
  
exposure_formatted <- exposure_SNPs %>% 
  format_data(
    type = "exposure",
    phenotype_col = "trait_name",
    snp_col = "rsid.x",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "oa",
    pval_col = "p",
    gene_col = "drug_target",
    min_pval = 1e-200,
    chr_col = "chr.x",
    pos_col = "bp.x"
  )
  
outcome_formatted <- outcome_SNPs %>% 
   format_data(
    type = "outcome",
    phenotype_col = "openGWAS_outcome",
    snp_col = "rsid",
    beta_col = "beta",
    se_col = "se",
    eaf_col = "eaf",
    effect_allele_col = "ea",
    other_allele_col = "nea",
    pval_col = "p",
    samplesize_col = "n",
    id_col = "id",
    min_pval = 1e-200,
    info_col = "indication",
    chr_col = "chr",
    pos_col = "position"
  )
  
# Harmonise
  
harmonised <- harmonise_data(exposure_formatted, outcome_formatted)

#######################################################
# 10. MR
#######################################################

mr_res <- mr(harmonised)

mr_res$indication <- ifelse(
  mr_res$outcome %in% opengwas_outcomes$openGWAS_outcome,
  TRUE,
  FALSE
)
```


```{r}
#| label: Warfarin_MR_Results
#| include: true
#| echo: false
#| fig-width: 10
#| fig-height: 10

# Bubble plot

ggplot(mr_res,
       aes(x = outcome,
           y = exposure,
           size = -log10(pval),
           colour = indication)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_size_continuous(name = "-log10(P)") +
  labs(
    x = "Outcome",
    y = "Exposure",
    colour = "Drug Indication"
  )
```


```{r}
#| eval: false
#| include: false

quarto::quarto_render(
  "scripts/06_Variant_Outcome_Associations/065_MR_Plotting.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```