---
title: "MR Results - Drug Bank Primary Indications and openGWAS outcomes"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
        fig-width: 8
        fig-height: 10
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)
library(tidyr)
library(ieugwasr)
library(TwoSampleMR)
library(ggplot2)

# Read environment file

readRenviron(here("config.env"))

interim_dir <- Sys.getenv("interimdatadir")
processed_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

exposure_file <- file.path(interim_dir, "predicted_outcomes/Primary_Indication_Mapping/MR_input/exposure_associations.csv")
outcome_file <- file.path(interim_dir, "predicted_outcomes/Primary_Indication_Mapping/MR_input/outcome_associations.csv")
targets_file <- file.path(interim_dir, "predicted_outcomes/Drug_Bank_targets/Drug_Bank_targets.csv")
indications_file <- file.path(interim_dir, "reported_outcomes/Drug_Bank_outcomes/Drug_Bank_Indications.csv")
MR_res_dir <- file.path(processed_dir, "Primary_indications_MR_Results")

```


```{r}
#| label: Reading in data
#| include: false

# Target of each drug
drug_targets <- fread(targets_file) %>% 
  filter(Primary == "Primary")

# Indication of each drug
drug_indications <- fread(indications_file) %>% 
  filter(`Approval Level` == "Prescription")

# Normalise some drug naming
drug_indications[which(drug_indications$Drug == "Valproate"), "Drug"] <- "Valproic acid"

# Drug target assocation data
exposure_data <- fread(exposure_file)

# Outcome association data
outcome_data <- fread(outcome_file) %>% 
  select("indication", "openGWAS_outcome", "id", "chr", "position", "rsid", 
         "ea", "nea", "eaf", "beta", "se", "p", "n")

MR_res_files <- list.files(MR_res_dir, full.names = TRUE)

```

# Drugs of Interest:
    * Captopril
    * Methotrexate
    * Mycophenolate mofetil
    * Thalidomide
    * Tretinoin
    * Valproate
    * Warfarin

# Captopril

## Primary Indications
```{r}
#| label: Captopril_Primary_Indications
#| include: true
#| echo: false

drug <- "Captopril"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Captopril Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Captopril openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Captopril_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Hypertension", "Heart failure", "Diabetic nephropathy", "Heart failure"),
  opengwas_id = c("ieu-b-5144", "ebi-a-GCST009541", "ebi-a-GCST90018832", "ebi-a-GCST009541"),
  sample_size = c("462,826", "977,323", "452,280", "977,323"),
  ncases = c("133,680", "47,309", "1,032", "47,309"),
  outcome_type = c("Binary", "Binary", "Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Captopril Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Captopril_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Captopril Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis Results
```{r}
#| label: Captopril_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Captopril", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Captopril_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Methotrexate

## Primary Indications
```{r}
#| label: Methotrexate_Primary_Indications
#| include: true
#| echo: false

drug <- "Methotrexate"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Methotrexate Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Methotrexate openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Methotrexate_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Psoriasis vulgaris", "Mature T/NK-cell lymphomas", "Rheumatoid arthritis",
                       "Juvenile rheumatoid arthritis", "Psoriasis vulgaris", "Head and neck cancer",
                       "Malignant lymphoma", "Meningitis"),
  opengwas_id = c("ebi-a-GCST90018907", "finn-b-CD2_TNK_LYMPHOMA", "ebi-a-GCST90018910",
                  "ebi-a-GCST90018873", "ebi-a-GCST90018907", "ieu-b-4912", "ebi-a-GCST90018878",
                  "finn-b-MENINGITIS"),
  sample_size = c("483174", "150 (ncase)", "417256", "409217", "483174", "373122", "490803",
                  "770 (ncase)"),
  outcome_type = c("Binary", "Binary", "Binary", "Binary", "Binary", "Binary", "Binary", "Binary")
)

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Psoriasis vulgaris", "Mature T/NK-cell lymphomas", "Rheumatoid arthritis",
                       "Juvenile rheumatoid arthritis", "Psoriasis vulgaris", "Head and neck cancer",
                       "Malignant lymphoma", "Meningitis"),
  opengwas_id = c("ebi-a-GCST90018907", "finn-b-CD2_TNK_LYMPHOMA", "ebi-a-GCST90018910",
                  "ebi-a-GCST90018873", "ebi-a-GCST90018907", "ieu-b-4912", "ebi-a-GCST90018878",
                  "finn-b-MENINGITIS"),
  sample_size = c("483,174", "NA", "417,256", "409,217", "483,174", "373,122", "490,803",
                  "NA"),
  ncases = c("5,072", "150", "8,255", "216", "5,072", "1,106", "3,546", "770"),
  outcome_type = c("Binary", "Binary", "Binary", "Binary", "Binary", "Binary", "Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Methotrexate Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Methotrexate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Methotrexate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results
```{r}
#| label: Methotrexate_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Methotrexate", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Methotrexate_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme(axis.text.y = element_text(size = 6)) +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

# Mycophenolate mofetil

## Primary Indications
According to DrugBank, Mycophenolate mofetil is used for the following indications:

 - Transplant organ rejection

However, it is reported as being used as Adjunct therapy, and has therefore been filtered out as not the primary drug used for this indication.
Therefore, there are no primary indications for this drug, and no associated outcomes in openGWAS. 


# Thalidomide

## Primary Indications

```{r}
#| label: Thalidomide_Primary_Indications
#| include: true
#| echo: false

drug <- "Thalidomide"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Thalidomide Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Thalidomide openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Thalidomide_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Erythema nodosum", "Erythema nodosum", "Erythema nodosum"),
  opengwas_id = c("finn-b-L12_ERYTHEMANODOSUM", "finn-b-L12_ERYTHEMANODOSUM", "finn-b-L12_ERYTHEMANODOSUM"),
  sample_size = c("NA", "NA", "NA"),
  ncases = c("433", "433", "433"),
  outcome_type = c("Binary", "Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Thalidomide Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Thalidomide_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Thalidomide Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis Results
```{r}
#| label: Thalidomide_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Thalidomide", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Thalidomide_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x") +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Tretinoin

## Primary Indications

```{r}
#| label: Tretinoin_Primary_Indications
#| include: true
#| echo: false

drug <- "Tretinoin"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Tretinoin Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Tretinoin openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Tretinoin_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Acne vulgaris", "Acute myeloid leukaemia"),
  opengwas_id = c("finn-b-L12_ACNE_VULGARIS", "finn-b-C3_AML"),
  sample_size = c("NA", "NA"),
  ncases = c("1092","90"),
  outcome_type = c("Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Tretinoin Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Tretinoin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Tretinoin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```


## MR Analysis Results
```{r}
#| label: Tretinoin_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Tretinoin", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Tretinoin_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x") +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Valproate

## Primary Indications

```{r}
#| label: Valproate_Primary_Indications
#| include: true
#| echo: false

drug <- "Valproic acid"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Valproate Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Valproate openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Valproate_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Epilepsy", "Manic episode", "Epilepsy", "Diagnoses - main ICD10: G43 Migraine"),
  opengwas_id = c("ebi-a-GCST90018840", "finn-b-F5_MANIA", "ebi-a-GCST90018840", "ukb-d-G43"),
  sample_size = c("458,310", "NA", "458,310", "361,194"),
  ncases = c("4,382", "631", "4,382", "1,072"),
  outcome_type = c("Binary", "Binary", "Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Valproate Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Valproate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Valproate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results
```{r}
#| label: Valproate_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Valproic acid", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Valproate_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Warfarin

## Primary Indications

```{r}
#| label: Warfarin_Primary_Indications
#| include: true
#| echo: false

drug <- "Warfarin"

# What are the indications
indications <- drug_indications %>% 
  filter(Drug == drug)

indication_list <- paste0("^", paste(indications$Indication, collapse = "$|^"), "$")

knitr::kable(
  indications,
  caption = "Warfarin Primary Indications",
)

opengwas_outcomes <- outcome_data
opengwas_outcomes$indication <- gsub("_", " ", opengwas_outcomes$indication)
opengwas_outcomes <- opengwas_outcomes[grep(indication_list, opengwas_outcomes$indication),] %>% 
  select(c("indication","openGWAS_outcome")) %>% 
  unique()

knitr::kable(
  opengwas_outcomes,
  caption = "Warfarin openGWAS Outcomes",
)

```

## Chosen OpenGWAS Outcomes
```{r}
#| label: Warfarin_Chosen_OpenGWAS_Outcomes
#| include: true
#| echo: false

studies <- data.table(
  indication = unique(indications$Indication),
  opengwas_outcome = c("Non-cancer illness code, self-reported: deep venous thrombosis (dvt)", "Venous thromboembolism",
                       "Pulmonary embolism"),
  opengwas_id = c("ukb-b-12040", "ukb-d-I9_VTE", "finn-b-I9_PULMEMB"),
  sample_size = c("462,933", "361,194", "NA"),
  ncases = c("9,241", "4,620", "4,185"),
  outcome_type = c("Binary", "Binary", "Binary")
)

knitr::kable(
  studies,
  caption = "Warfarin Chosen OpenGWAS Outcomes",
)

```

## Targets
```{r}
#| label: Warfarin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Warfarin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results
```{r}
#| label: Warfarin_MR
#| include: false
#| echo: false

chosen_outcomes <- paste0("^",
                    paste(unique(studies$opengwas_id), collapse = "$|^"),
                    "$")

mr_res <- MR_res_files[grep("Warfarin", MR_res_files)] %>% 
  fread()

mr_res_plot <- mr_res[grep(chosen_outcomes, mr_res$id.outcome),] %>% 
  as.data.frame()

plot_df <- mr_res_plot %>% 
  mutate(
    OR = exp(b),
    lower = exp(b - 1.96 * se),
    upper = exp(b + 1.96 * se),
    outcome = factor(openGWAS_outcome),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

```{r}
#| label: Warfarin_MR_Plot
#| include: true
#| echo: false
#| warning: false
#| message: false

ggplot(plot_df,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ openGWAS_outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


```{r}
#| eval: false
#| include: false

quarto::quarto_render(
  "scripts/06_Variant_Outcome_Associations/066_MR_Plotting.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```