---
title: "MR Results - MR PREG"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
        fig-width: 8
        fig-height: 10
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)
library(tidyr)
library(ieugwasr)
library(TwoSampleMR)
library(ggplot2)

# Read environment file

readRenviron(here("config.env"))

raw_dir <- Sys.getenv("rawdatadir")
interim_dir <- Sys.getenv("interimdatadir")
processed_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

targets_file <- file.path(interim_dir, "predicted_outcomes/Drug_Bank_targets/Drug_Bank_targets.csv")
MR_res_dir <- file.path(processed_dir, "MR_PREG_MR_Results")

MR_PREG_outcome_names_file <- file.path(raw_dir, "predicted_outcomes/MR_PREG/MR_PREG_Outcome_Names.csv")

```

```{r}
#| label: Reading in data
#| include: false

# Target of each drug
drug_targets <- fread(targets_file) %>% 
  filter(Primary == "Primary")

# Side effect of each drug - This will be the extracted side effects from reports

# Mapped Outcome data - This will include the MR PREG outcomes associated with the extracted side effects

# MR output
MR_res_files <- list.files(MR_res_dir, full.names = TRUE)

# MR PREG outcome names
MR_PREG_outcome_names <- fread(MR_PREG_outcome_names_file)

```

# Drugs of Interest:
    * Captopril
    * Methotrexate
    * Mycophenolate mofetil
    * Thalidomide
    * Tretinoin
    * Valproate
    * Warfarin

## To Include:
  * Side effects identified by report extraction
  * How these side effects map to MR PREG outcomes
  * Therefore which MR PREG outcomes are relevant to the drug of interest

# Captopril

## Side Effects
```{r}
#| label: Captopril_Side_Effects
#| include: true
#| echo: false

drug <- "Captopril"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Captopril Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Captopril MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Captopril_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Captopril Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Captopril_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Captopril", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Captopril_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Captopril_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Methotrexate

## Side Effects
```{r}
#| label: Methotrexate_Side_Effects
#| include: true
#| echo: false

drug <- "Methotrexate"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Methotrexate Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Methotrexate MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Methotrexate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Methotrexate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Methotrexate_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Methotrexate", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Methotrexate_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() + 

  theme(
    axis.text.y = element_text(size = 5)
  ) +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Methotrexate_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +

  theme(
    axis.text.y = element_text(size = 5)
  ) +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Mycophenolate mofetil

## Side Effects
```{r}
#| label: Mycophenolate_mofetil_Side_Effects
#| include: true
#| echo: false

drug <- "Mycophenolate mofetil"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Mycophenolate mofetil Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Mycophenolate mofetil MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Mycophenolate_mofetil_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Mycophenolate mofetil Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Mycophenolate_mofetil_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Mycophenolate_mofetil", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Mycophenolate_mofetil_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Mycophenolate_mofetil_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Thalidomide

## Side Effects
```{r}
#| label: Thalidomide_Side_Effects
#| include: true
#| echo: false

drug <- "Thalidomide"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Thalidomide Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Thalidomide MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Thalidomide_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Thalidomide Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Thalidomide_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Thalidomide", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Thalidomide_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Thalidomide_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Tretinoin

## Side Effects
```{r}
#| label: Tretinoin_Side_Effects
#| include: true
#| echo: false

drug <- "Tretinoin"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Tretinoin Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Tretinoin MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Tretinoin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Tretinoin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Tretinoin_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Tretinoin", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Tretinoin_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Tretinoin_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Valproate

## Side Effects
```{r}
#| label: Valproate_Side_Effects
#| include: true
#| echo: false

drug <- "Valproic acid"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Valproate Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Valproate MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Valproate_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Valproate Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Valproate_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Valproic_acid", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Valproate_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Valproate_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


# Warfarin

## Side Effects
```{r}
#| label: Warfarin_Side_Effects
#| include: true
#| echo: false

drug <- "Warfarin"

# What are the side effects


#knitr::kable(
#  side_effects,
#  caption = "Warfarin Side Effects",
#)

# What MR PREG outcomes are associated with these side effects

#knitr::kable(
#  mrpreg_outcomes,
#  caption = "Warfarin MR PREG Outcomes",
#)

```

## Targets
```{r}
#| label: Warfarin_Targets
#| include: true
#| echo: false


# What are the targets
targets <- drug_targets %>% 
  filter(Drug == drug) %>% 
  select(c("Target", "Gene_Name")) %>% 
  unique()

knitr::kable(
  targets,
  caption = "Warfarin Targets",
)

target_list <- paste0("^", paste(targets$Gene_Name, collapse = "$|^"), "$")
```

## MR Analysis Results

Once I have a good data frame of side effects for each drug extracted from the reports, I can select/highlight the relevant MR PREG outcomes for each drug.

```{r}
#| label: Warfarin_MR
#| include: false
#| echo: false

mr_res <- MR_res_files[grep("Warfarin", MR_res_files)] %>% 
  fread()

mr_res <- left_join(mr_res, MR_PREG_outcome_names, by = c("id.outcome" = "MR_PREG_Outcome"))

mr_res$binary <- ifelse(!is.na(mr_res$ncase.outcome), "Yes", "No")

plot_df <- mr_res %>%
  mutate(
    OR = if_else(binary == "Yes", exp(b), NA_real_),
    lower = if_else(binary == "Yes",
                    exp(b - 1.96 * se),
                    b - 1.96 * se),
    upper = if_else(binary == "Yes",
                    exp(b + 1.96 * se),
                    b + 1.96 * se),
    outcome = factor(Meaning),
    sig = pval < 0.05
  )

genes <- unique(plot_df$gene.exposure)

gene_cols <- setNames(
  scales::hue_pal()(length(genes)),
  genes
)

plot_df$fill_col <- ifelse(
  plot_df$sig,
  gene_cols[plot_df$gene.exposure],
  "white"
)
```

### Binary Outcomes
```{r}
#| label: Warfarin_MR_Plot_Binary
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 100

plot_df_bin <- plot_df %>% filter(binary == "Yes")

ggplot(plot_df_bin,
       aes(
         y = id.exposure,
         x = OR,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_bin$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 1, linetype = 2) +
  
  scale_x_log10() +
  
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Odds Ratio per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```

### Continuous Outcomes
```{r}
#| label: Warfarin_MR_Plot_Continuous
#| include: true
#| echo: false
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 30

plot_df_cont <- plot_df %>% filter(binary == "No")

ggplot(plot_df_cont,
       aes(
         y = id.exposure,
         x = b,
         xmin = lower,
         xmax = upper
       )) +
  
  geom_errorbarh(height = 0.2,
                 aes(colour = gene.exposure)) +
                 
  geom_point(
    aes(colour = gene.exposure),
    shape = 21,
    fill = plot_df_cont$fill_col,
    size = 2
  ) +

  geom_vline(xintercept = 0, linetype = 2) +
    
  facet_wrap(~ outcome, scales = "free_x", ncol = 2) +

  scale_colour_manual(values = gene_cols) +
  
  theme_bw() +
  
  labs(x = "Effect Size per 1 SD increase in QTL", y = "Drug Target Measure", colour = "Drug Target")

```


```{r}
#| eval: false
#| include: false

quarto::quarto_render(
  "scripts/06_Variant_Outcome_Associations/068_MR_Preg_Plotting.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```