---
title: "Evaluating Ontology Mapping using Multiple Language Models"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(here)
library(gt)
library(kableExtra)

# Read environment file
#load_dot_env("config.env")
readRenviron(here("config.env"))

interim_dir <- Sys.getenv("interimdatadir")
processed_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

# Mapping outputs
mapping_output_dir <- file.path(interim_dir, "ontology_mapping/output_data")
pipeline_output_dir <- file.path(processed_dir, "ONE_TEST_test_outcomes/test/two_test")

# Manually curated expected cosine similarity values
expected_path <- file.path(interim_dir, "ontology_mapping/input_data/test/test_all.csv")
control_outcomes <- fread(expected_path)

```

# Mapping Pipeline

Ontology mapping was used to compare phenotypic outcomes from different sources using different terminologies.

The different sources included:

  * Observed/reported outcomes:
    * FAERS
      * All outcomes
      * Congenital, Familial and Genetic disorders
      * Pregnancy, Puerperium and Perinatal conditions
    * UKTIS Monographs
    * Reprotox
    * TERIS
    * SmPCs
  * Genetic databases used for prediction:
    * OMIM
    * Genotype-Phenotype Map
    * IEU OpenGWAS
    * MR PREG
    * 100,000 Genomes Project

The pipeline used the following steps:

  1. Biobert models 

    a. Generate embeddings for each outcome term
    b. Calculate cosine similarity between each pair of outcomes
    c. Extract the 30 most similar observed outcome for each predicted outcome
  2. Qwen 

    a. Assess the similarity of the top 30 pairs of outcomes and generate a similarity score
    b. Only keep pairs with some association (score > 0)
  3. DeepSeek

    a. Assess the similarity of the outcomes output by Qwen and generate a similarity score
    b. Only accept pairs with some relationship (score > 1)

# Evaluating Mapping performance

This was done using known pairs of outcomes (n = 100):  
  * Some were similar with different names  
  * Some were different with similar names  

The number of pairs successfully mapped by each model was compared to the expected number of pairs.

## Assessing Biobert Model Performance

This function uses a Spearman rank correlation, it has:  
  * More focus on the order/ranking rather than the exact values  
  * Robust to the fact that cosine similarities may be scaled differently across models  

```{r}
#| label: Model Assessment Function
#| include: true
#| echo: false

model_assessment <- function(model){

  # Biobert results
  results_path <- file.path(mapping_output_dir, model, "test/test_one_test_two_test_similarity_matrix.csv")
  test_sim_matrix <- as.matrix(fread(results_path), rownames = 1)
  
  test_observed <- data.frame(
    terms_one = rownames(test_sim_matrix),
    terms_two = colnames(test_sim_matrix),
    cos_sim = diag(test_sim_matrix)
  )

  # Upper and lower bounds of expected cosine similarity
  test_expected <- control_outcomes %>%
    select(One_Outcome, Two_Outcome, Control, Expected_Cosine_Range)

  test_expected$lower_expected <- NA
  test_expected$upper_expected <- NA
  
  for (i in 1:nrow(test_expected)){
    test_expected$lower_expected[i] <- unlist(strsplit(test_expected$Expected_Cosine_Range[i], split = "-"))[1]
    test_expected$upper_expected[i] <- unlist(strsplit(test_expected$Expected_Cosine_Range[i], split = "-"))[2]
  }
  
  # Combine observed and expected data for comparison
  test_comparison <- left_join(test_observed, test_expected, by = c("terms_one" = "One_Outcome", "terms_two" = "Two_Outcome"))
  
  predicted_sim <- test_comparison$cos_sim
  lower_bound <- as.numeric(test_comparison$lower_expected)
  upper_bound <- as.numeric(test_comparison$upper_expected)
  
  cor_upper <- cor(predicted_sim, upper_bound, method = "spearman")
  cor_lower <- cor(predicted_sim, lower_bound, method = "spearman")
  
  return(list(upper_limit = cor_upper, lower_limit = cor_lower))
}

```

### Model: all-MiniLM-L6-v2
```{r}
#| label: all-MiniLM-L6-v2 assessmenet
#| include: true
#| echo: false

model <- "all-MiniLM-L6-v2"

all_MiniLM_L6_v2_results <- model_assessment(model = model)

print(paste(model, "has the following Spearman's rank outputs for upper and lower expected bound:"))

all_MiniLM_L6_v2_results
  
```

### Model: all-mpnet-base-v2
```{r}
#| label: all-mpnet-base-v2 assessment
#| include: true
#| echo: false

model <- "all-mpnet-base-v2"

all_mpnet_base_v2_results <- model_assessment(model = model)

print(paste(model, "has the following Spearman's rank outputs for upper and lower expected bound:"))

all_mpnet_base_v2_results

```

### Model: kamalkraj/BioSimCSE-BioLinkBERT-BASE
```{r}
#| label: kamalkraj/BioSimCSE-BioLinkBERT-BASE assessment
#| include: true
#| echo: false

model <- "kamalkraj/BioSimCSE-BioLinkBERT-BASE"

BioSimCSE_BioLinkBERT_BASE_results <- model_assessment(model = model)

print(paste(model, "has the following Spearman's rank outputs for upper and lower expected bound:"))

BioSimCSE_BioLinkBERT_BASE_results

```

### Model: pritamdeka/S-PubMedBert-MS-MARCO
```{r}
#| label: pritamdeka/S-PubMedBert-MS-MARCO assessment
#| include: true
#| echo: false

model <- "pritamdeka/S-PubMedBert-MS-MARCO"

S_PubMedBert_MS_MARCO_results <- model_assessment(model = model)

print(paste(model, "has the following Spearman's rank outputs for upper and lower expected bound:"))

S_PubMedBert_MS_MARCO_results

```

### Model: UMCU/SapBERT-from-PubMedBERT-fulltext_bf16
```{r}
#| label: UMCU/SapBERT-from-PubMedBERT-fulltext_bf16 assessment
#| include: true
#| echo: false

model <- "UMCU/SapBERT-from-PubMedBERT-fulltext_bf16"

SapBERT_from_PubMedBERT_fulltext_bf16_results <- model_assessment(model = model)

print(paste(model, "has the following Spearman's rank outputs for upper and lower expected bound:"))

SapBERT_from_PubMedBERT_fulltext_bf16_results  
```

Given these results, the model with the highest Spearman's rank correlation for both upper and lower expected bounds was considered the best performing model for ontology mapping in this context.

Chosen Biobert model: `UMCU/SapBERT-from-PubMedBERT-fulltext_bf16`

This was used for the rest of the pipeline.

## Assessing Pipeline Performance

### Qwen performance

```{r}
#| label: qwen assessment
#| include: true
#| echo: false

pipeline_output <- file.path(pipeline_output_dir, "test_one_test_two_test_full.csv") %>%
  fread()

test_info <- control_outcomes %>%
  select(One_Outcome, Two_Outcome, Control, Relationship_Type)

output_comparison <- full_join(pipeline_output, test_info, by = c("Predicted_term" = "One_Outcome", "Observed_term" = "Two_Outcome"))

true_positives <- output_comparison %>%
  filter(Control == "Positive") %>%
  filter(!is.na(Confidence_Qwen))

true_negatives <- output_comparison %>%
  filter(Control == "Negative") %>%
  filter(is.na(Confidence_Qwen))

false_positives <- output_comparison %>%
  filter(Control == "Negative") %>%
  filter(!is.na(Confidence_Qwen))

false_negatives <- output_comparison %>%
  filter(Control == "Positive") %>%
  filter(is.na(Confidence_Qwen))

```

```{r}
#| label: Qwen Confusion Matrix
#| include: true
#| echo: false

TP <- nrow(true_positives)
FP <- nrow(false_positives)
FN <- nrow(false_negatives)
TN <- nrow(true_negatives)

confusion_df <- data.frame(
  Prediction = c("Positive", "Negative"),
  Positive = c(TP, FN),
  Negative = c(FP, TN)
)

confusion_df %>%
  gt() %>%
  tab_spanner(
    label = "Observation",
    columns = c(Positive, Negative)
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "right",
      color = "light grey",
      weight = px(1.2)
    ),
    locations = list(
      cells_body(columns = Prediction),
      cells_column_labels(columns = Prediction)
    )
  )  

```

### Performance Metrics

```{r}
#| label: Qwen Performance Metrics
#| include: true
#| echo: false

accuracy    <- (TP + TN) / (TP + FP + FN + TN)
sensitivity <- TP / (TP + FN)
specificity <- TN / (TN + FP)
precision   <- TP / (TP + FP)
npv         <- TN / (TN + FN)
f1          <- 2 * precision * sensitivity / (precision + sensitivity)

cat(paste("The proportion of all predictions that were correct.\nAccuracy:", round(accuracy, 2)))
cat(paste("Of all the truly positive observations, how many did the model correctly identify?\nSensitivity:", round(sensitivity, 2)))
cat(paste("Of all the truly negative observations, how many did the model correctly identify?\nSpecificity:", round(specificity, 2)))
cat(paste("Of all the predictions labelled positive, how many were actually positive?\nPrecision:", round(precision, 2)))
cat(paste("Of all the predictions labelled negative, how many were actually negative?\nNPV:", round(npv, 2)))
cat(paste("A single measure that balances precision and sensitivity.\nF1 Score:", round(f1, 2)))

```

### Whole pipeline performance

```{r}
#| label: pipeline assessment
#| include: true
#| echo: false

true_positives <- output_comparison %>%
  filter(Control == "Positive") %>%
  filter(Decision_Deepseek == "ACCEPT")

true_negatives <- output_comparison %>%
  filter(Control == "Negative") %>%
  filter(Decision_Deepseek != "ACCEPT")

false_positives <- output_comparison %>%
  filter(Control == "Negative") %>%
  filter(Decision_Deepseek == "ACCEPT")

false_negatives <- output_comparison %>%
  filter(Control == "Positive") %>%
  filter(Decision_Deepseek != "ACCEPT")

```

```{r}
#| label: Confusion Matrix
#| include: true
#| echo: false

TP <- nrow(true_positives)
FP <- nrow(false_positives)
FN <- nrow(false_negatives)
TN <- nrow(true_negatives)

confusion_df <- data.frame(
  Prediction = c("Positive", "Negative"),
  Positive = c(TP, FN),
  Negative = c(FP, TN)
)

confusion_df %>%
  gt() %>%
  tab_spanner(
    label = "Observation",
    columns = c(Positive, Negative)
  ) %>%
  tab_style(
    style = cell_borders(
      sides = "right",
      color = "light grey",
      weight = px(1.2)
    ),
    locations = list(
      cells_body(columns = Prediction),
      cells_column_labels(columns = Prediction)
    )
  )  

```

### Performance Metrics

```{r}
#| label: Performance Metrics
#| include: true
#| echo: false

accuracy    <- (TP + TN) / (TP + FP + FN + TN)
sensitivity <- TP / (TP + FN)
specificity <- TN / (TN + FP)
precision   <- TP / (TP + FP)
npv         <- TN / (TN + FN)
f1          <- 2 * precision * sensitivity / (precision + sensitivity)

cat(paste("The proportion of all predictions that were correct.\nAccuracy:", round(accuracy, 2)))
cat(paste("Of all the truly positive observations, how many did the model correctly identify?\nSensitivity:", round(sensitivity, 2)))
cat(paste("Of all the truly negative observations, how many did the model correctly identify?\nSpecificity:", round(specificity, 2)))
cat(paste("Of all the predictions labelled positive, how many were actually positive?\nPrecision:", round(precision, 2)))
cat(paste("Of all the predictions labelled negative, how many were actually negative?\nNPV:", round(npv, 2)))
cat(paste("A single measure that balances precision and sensitivity.\nF1 Score:", round(f1, 2)))

```

### Taking a closer look at the false positives and false negatives

*False Positives*

The model has predicted that the following unrelated pairs of outcomes are similar.

```{r}
#| label: False Positives
#| include: true
#| echo: false

kable(false_positives) %>%
  kable_styling() 

```


*False Negatives*

The model has failed to predict that the following related pairs of outcomes are similar.

```{r}
#| label: False Negatives
#| include: true
#| echo: false

kable(false_negatives) %>%
  kable_styling() 

```



```{r}
#| eval: false
#| include: false

quarto::quarto_render(
  "scripts/07_Ontology_Mapping/075d_Mapping_Evaluation.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```