---
title: "Presenting Collated Ontology Terms - Genotype-Phenotype Map and FAERS Outcomes"
author: "Alice Robinson"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: 
    html:
        toc: true
        toc-location: left
        embed-resources: true
---

```{r}
#| label: setup
#| include: false

library(dotenv)
library(data.table)
library(tidyverse)
library(jsonlite)
library(dplyr)
library(here)
library(knitr)
library(kableExtra)

# Read environment file
#load_dot_env("config.env")
readRenviron(here("config.env"))

interim_dir <- Sys.getenv("interimdatadir")
base_dir <- Sys.getenv("processeddatadir")
resultsdir <- Sys.getenv("docsdir")

LLM_results_dir <- file.path(base_dir, "GPMAP_FAERS_outcomes")
term_counts <- file.path(interim_dir, "ontology_mapping/input_data")

```

# Mapped Predicted and Observed Drug outcomes

**Predicted outcomes** - Genotype-Phenotype Map traits colocalising with genetic variants associated with drug target gene

**Observed outcomes** - FAERS adverse events associated with drugs. This dataset includes all outcomes and those that are pregnancy and congenitally related.

## The terms have been mapped in three steps:
    1. Biobert model to calculate cosine similarity between GPMap and FAERS terms. 
     -> Filtered for the top 30 most similar FAERS terms for each GPMap term.
    2. Qwen2.5 model to generate a confidence score of 0-5 for each GPMap-FAERS term pair based on the closeness of the traits. 
     -> Those with a score of 0 were filtered out.
    3. DeepSeek model to generate a confidence score of 0-5 for each GPMap-FAERS term pair based on the closeness of the traits. 
     -> Those with a score 0-1 were rejected.

## Confidence Score Interpretation:
    5: Exact or near-perfect conceptual match
    4: Strong pathophysiological relationship (same disease category, direct complications)
    3: Moderate relationship (shared mechanisms, risk factors, or related conditions)
    2: Weak relationship (indirect connections, shared symptoms)
    1: Very loose association
    0: No association

# Drugs of Interest:
    * Captopril
    * Methotrexate
    * Mycophenolate mofetil
    * Thalidomide
    * Tretinoin
    * Valproate
    * Warfarin

# Captopril

```{r}
#| label: Captopril term counts
#| include: true
#| echo: false

captopril_count_gpmap <- nrow(fread(file.path(term_counts, "captopril/captopril_gpmap.csv")))
captopril_count_faers_all <- nrow(fread(file.path(term_counts, "captopril/captopril_faers_all.csv")))
captopril_count_faers_cong <- nrow(fread(file.path(term_counts, "captopril/captopril_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", captopril_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", captopril_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", captopril_count_faers_cong), "\n")

```

## All outcomes

```{r}
#| label: Captopril all
#| include: true
#| echo: false

captopril_results_path <- file.path(LLM_results_dir, "captopril/all/captopril_gpmap_faers_LLM_all_accepted.csv")

captopril_results <- fread(captopril_results_path) %>%
    select(Predicted_term, Observed_term)

kable(captopril_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Captopril congenital
#| include: true
#| echo: false

captopril_results_path <- file.path(LLM_results_dir, "captopril/cong/captopril_gpmap_faers_LLM_cong_accepted.csv")

captopril_results <- fread(captopril_results_path) %>%
    select(Predicted_term, Observed_term)

kable(captopril_results) %>%
  kable_styling() 

```

# Methotrexate

```{r}
#| label: Methotrexate term counts
#| include: true
#| echo: false

methotrexate_count_gpmap <- nrow(fread(file.path(term_counts, "methotrexate/methotrexate_gpmap.csv")))
methotrexate_count_faers_all <- nrow(fread(file.path(term_counts, "methotrexate/methotrexate_faers_all.csv")))
methotrexate_count_faers_cong <- nrow(fread(file.path(term_counts, "methotrexate/methotrexate_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", methotrexate_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", methotrexate_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", methotrexate_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Methotrexate all
#| include: true
#| echo: false

methotrexate_results_path <- file.path(LLM_results_dir, "methotrexate/all/methotrexate_gpmap_faers_LLM_all_accepted.csv")

methotrexate_results <- fread(methotrexate_results_path) %>%
    select(Predicted_term, Observed_term)

kable(methotrexate_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only
```{r}
#| label: Methotrexate cong
#| include: true
#| echo: false

methotrexate_results_path <- file.path(LLM_results_dir, "methotrexate/cong/methotrexate_gpmap_faers_LLM_cong_accepted.csv")

methotrexate_results <- fread(methotrexate_results_path) %>%
    select(Predicted_term, Observed_term)

kable(methotrexate_results) %>%
  kable_styling() 

```

# Mycophenolate mofetil

```{r}
#| label: Mycophenolate mofetil term counts
#| include: true
#| echo: false

mycophenolate_mofetil_count_gpmap <- nrow(fread(file.path(term_counts, "mycophenolate_mofetil/mycophenolate_mofetil_gpmap.csv")))
mycophenolate_mofetil_count_faers_all <- nrow(fread(file.path(term_counts, "mycophenolate_mofetil/mycophenolate_mofetil_faers_all.csv")))
mycophenolate_mofetil_count_faers_cong <- nrow(fread(file.path(term_counts, "mycophenolate_mofetil/mycophenolate_mofetil_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", mycophenolate_mofetil_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", mycophenolate_mofetil_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", mycophenolate_mofetil_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Mycophenolate mofetil all
#| include: true
#| echo: false

mycophenolate_mofetil_results_path <- file.path(LLM_results_dir, "mycophenolate_mofetil/all/mycophenolate_mofetil_gpmap_faers_LLM_all_accepted.csv")

mycophenolate_mofetil_results <- fread(mycophenolate_mofetil_results_path) %>%
    select(Predicted_term, Observed_term)

kable(mycophenolate_mofetil_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Mycophenolate mofetil cong
#| include: true
#| echo: false

mycophenolate_mofetil_results_path <- file.path(LLM_results_dir, "mycophenolate_mofetil/cong/mycophenolate_mofetil_gpmap_faers_LLM_cong_accepted.csv")

mycophenolate_mofetil_results <- fread(mycophenolate_mofetil_results_path) %>%
    select(Predicted_term, Observed_term)

kable(mycophenolate_mofetil_results) %>%
  kable_styling() 

```

# Thalidomide

```{r}
#| label: Thalidomide term counts
#| include: true
#| echo: false

thalidomide_count_gpmap <- nrow(fread(file.path(term_counts, "thalidomide/thalidomide_gpmap.csv")))
thalidomide_count_faers_all <- nrow(fread(file.path(term_counts, "thalidomide/thalidomide_faers_all.csv")))
thalidomide_count_faers_cong <- nrow(fread(file.path(term_counts, "thalidomide/thalidomide_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", thalidomide_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", thalidomide_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", thalidomide_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Thalidomide all
#| include: true
#| echo: false

thalidomide_results_path <- file.path(LLM_results_dir, "thalidomide/all/thalidomide_gpmap_faers_LLM_all_accepted.csv")

thalidomide_results <- fread(thalidomide_results_path) %>%
     select(Predicted_term, Observed_term)

kable(thalidomide_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Thalidomide cong
#| include: true
#| echo: false

thalidomide_results_path <- file.path(LLM_results_dir, "thalidomide/cong/thalidomide_gpmap_faers_LLM_cong_accepted.csv")

thalidomide_results <- fread(thalidomide_results_path) %>%
     select(Predicted_term, Observed_term)

kable(thalidomide_results) %>%
  kable_styling() 

```

# Tretinoin

```{r}
#| label: Tretinoin term counts
#| include: true
#| echo: false

tretinoin_count_gpmap <- nrow(fread(file.path(term_counts, "tretinoin/tretinoin_gpmap.csv")))
tretinoin_count_faers_all <- nrow(fread(file.path(term_counts, "tretinoin/tretinoin_faers_all.csv")))
tretinoin_count_faers_cong <- nrow(fread(file.path(term_counts, "tretinoin/tretinoin_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", tretinoin_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", tretinoin_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", tretinoin_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Tretinoin all
#| include: true
#| echo: false

tretinoin_results_path <- file.path(LLM_results_dir, "tretinoin/all/tretinoin_gpmap_faers_LLM_all_accepted.csv")

tretinoin_results <- fread(tretinoin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(tretinoin_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Tretinoin cong
#| include: true
#| echo: false

tretinoin_results_path <- file.path(LLM_results_dir, "tretinoin/cong/tretinoin_gpmap_faers_LLM_cong_accepted.csv")

tretinoin_results <- fread(tretinoin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(tretinoin_results) %>%
  kable_styling() 

```

# Valproate

```{r}
#| label: Valproate term counts
#| include: true
#| echo: false

valproate_count_gpmap <- nrow(fread(file.path(term_counts, "valproate/valproate_gpmap.csv")))
valproate_count_faers_all <- nrow(fread(file.path(term_counts, "valproate/valproate_faers_all.csv")))
valproate_count_faers_cong <- nrow(fread(file.path(term_counts, "valproate/valproate_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", valproate_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", valproate_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", valproate_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Valproate all
#| include: true
#| echo: false

valproate_results_path <- file.path(LLM_results_dir, "valproate/all/valproate_gpmap_faers_LLM_all_accepted.csv")

valproate_results <- fread(valproate_results_path) %>%
     select(Predicted_term, Observed_term)

kable(valproate_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Valproate cong
#| include: true
#| echo: false

valproate_results_path <- file.path(LLM_results_dir, "valproate/cong/valproate_gpmap_faers_LLM_cong_accepted.csv")

valproate_results <- fread(valproate_results_path) %>%
     select(Predicted_term, Observed_term)

kable(valproate_results) %>%
  kable_styling() 

```

# Warfarin

```{r}
#| label: Warfarin term counts
#| include: true
#| echo: false

warfarin_count_gpmap <- nrow(fread(file.path(term_counts, "warfarin/warfarin_gpmap.csv")))
warfarin_count_faers_all <- nrow(fread(file.path(term_counts, "warfarin/warfarin_faers_all.csv")))
warfarin_count_faers_cong <- nrow(fread(file.path(term_counts, "warfarin/warfarin_faers_cong.csv")))

cat(paste("Number of Genotype-Phenotype Map Traits:", warfarin_count_gpmap), "\n")
cat(paste("Number of FAERS Reported Outcomes (all):", warfarin_count_faers_all), "\n")
cat(paste("Number of FAERS Reported Outcomes (congenital):", warfarin_count_faers_cong), "\n")

```


## All outcomes

```{r}
#| label: Warfarin all
#| include: true
#| echo: false

warfarin_results_path <- file.path(LLM_results_dir, "warfarin/all/warfarin_gpmap_faers_LLM_all_accepted.csv")

warfarin_results <- fread(warfarin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(warfarin_results) %>%
  kable_styling() 

```

## Pregnancy and congenitally related outcomes only

```{r}
#| label: Warfarin cong
#| include: true
#| echo: false

warfarin_results_path <- file.path(LLM_results_dir, "warfarin/cong/warfarin_gpmap_faers_LLM_cong_accepted.csv")

warfarin_results <- fread(warfarin_results_path) %>%
     select(Predicted_term, Observed_term)

kable(warfarin_results) %>%
  kable_styling() 

```

# Render
```{r}
#| eval: false

quarto::quarto_render(
  "scripts/06_Ontology_Mapping/065c_GPMap_Traits_Mapping_Results_Collation.qmd",
  quarto_args = c("--output-dir", resultsdir)
)
```